The Natural History of Child Signals of Need in Utila, Honduras: An Exploratory Study

### 2024-10-14 Paper.qmd
Added dataframes for glmnet models. Minor fixes for knitting.
Paper: edited discussion.
Rest: minor changes.

The survey also included many demographic questions about the caretakers and their households. Those used in this study included the (1) ages of those in the household, (2) sexes of those in the household, (3) partnership status of the primary caregiver, (4) number of adults who provide childcare within the household, (5) highest level of education of the primary caregiver, (5) immigration status (in regards to Utila), and (6) monthly household income. These data allowed us to calculate the number of children (total, younger, and older) and adults within the household.

@@

@@

@@

@@

@@

>>>>

The survey also included many demographic questions about the caretakers and their households. Those used in this study included the (1) ages of those in the household, (2) sexes of those in the household, (3) partnership status of the primary caregiver, (4) number of adults who provide childcare within the household, (5) highest level of education of the primary caregiver, (5) immigration status (in regards to Utila), and (6) monthly household income. These data allowed us to calculate the number of children (total, younger, and older) and adults within the household. We also imputed data for caregiver age by setting the missing value to the mean. For the data used in analyses, this amounted to imputing data for one mother which affected two children.

@@

As seen in Western cultures [@zeifman2013; @potegal2003; @koch2003] (check for others), the frequencies of crying and temper tantrums decreased with age. One possibility is that older children have greater ability to satisfy one's desires and behave in ways which match cultural expectations of teenagers and adults. Another is that since one's reproductive value increases as one survives periods of high mortality risk in childhood, the intensity of conflicts of interest with parents decreases since these children are likely to reach adulthood and require less investment to reach adulthood than younger children, current or future [@daly1995]. For this reason, less costly signals may be more likely to work to increase investment, all else being equal\[\[4\]\]. That said, all else may often not be equal if investment in older children is costlier, something which may be reflected in caretakers seeing older children as relatively needy. The decrease in crying and tantrum frequencies with age may also result from the potential for greater reputational costs, something which may also increase the credibility of such signals among older children as previously mentioned.

There was also a negative association between child sadness and age. However, interpretation of this finding is more difficult due to the fact that, unlike crying and temper tantrums, sadness is thought to have important non-signaling functions including drawing one’s attention to a problem, motivating deeper analysis of one’s situation, and causing withdrawal from situations with low fitness prospects [@nesse2019; @nettle2009; @allen2003; @hagen2011]. By asking about sadness as a whole, our measure was unable to disentangle information caretakers gleaned from cues from information that came from signals. 
- difference in sadness or a difference in facial sadness?

@@

Consistent with signaling models which emphasize the role of conflicts of interest in favoring the use of costlier signals, conflict and signaling behavior were highly correlated. Although conflicts of interest do not always lead to behavioral conflict [@altmann1980; @bateson1994], it is likely many, if not most, conflicts are ultimately caused by conflicts of interest, potentially making it a useful proxy for the underlying conflicts of interest that may increase the benefits of deception and the costs of being deceived\[\[5\]\]. If this is the case, the correlation between signaling and conflict may represent children's attempts at credibly signaling their value of additional investment or the costliness of punishment [@syme2019] to parents who would have good reason to doubt the trustworthiness of simple verbal requests (or the trustworthiness of facial sadness and crying for those resorting to temper tantrums). However, it is also likely that parents view aversive forms of child signaling as conflict in and of itself.

@@

We found evidence of an interaction between how often a child cared for their siblings and the sex of the child in our models examining signal frequencies and cost, whereby there was a positive relationship between the frequency and summed costliness of child signals of need and alloparenting effort for girls but no relationship or a slightly negative one for boys. In many societies, girls are often expected to do more alloparenting within the household and often end up doing so [@helfrecht2016 @weisner1977; @barry1971; @turke1988]. Our models predict that girls in this sample are more likely to provide allocare (confirm), something which matches findins from focus groups of Utilian mothers. If this difference was the outcome of caretakers pressuring girls to do more or resulted in the expectation of more alloparenting, the increase in signaling with alloparenting effort seen among girls may be the result of a strategy to signal the costliness of their efforts in order to invest more in romantic relationships and status seeking behaviors outside the household.

@@

However, child alloparenting effort was negatively associated with caregiver reports of child conflict. One potential reason for this is since parents presumably value the child alloparenting within the family, they see their relationships with children who invest more in the household as less conflictual. In line with this reasoning, we found evidence of a child alloparenting effort by child age interaction, with our model predicting increased levels of conflict for older children with greater alloparenting. Ideally, we could control for signaling behavior to test this. However, the bidirectional relationship between child signaling and conflict with caretakers prevents this (explain better). It is also possible that our hypothesis about conflict over investment in the household is incorrect.

@@

### OLD Discussion

#### OLD Limitations
